### 2016-04-09 03-analytics.tex

\chapter{Data Analytics}

Hypotheses:

1. Skedge's differences from and additions to CDCS are usable and have real need

2. Skedge’s navigations-per-add and other metrics demonstrate effectiveness of the use cases
a) direct searching, and
b) course browsing

3. Skedge’s DSL is user-friendly; users learn more advanced search types over time by using it

\section{Usage}

\subsection{General}

Since November 3rd 2015 (137 days)
3,768 unique users
4,500 schedules
Average 90 sessions/day
Average 4.92 pages/session
Average 5:31 minutes/session
28\% of sessions are from new users

MOBILE RESULT

\subsection{Search}

% figs

  \subsubsection{Empty searches}

  Can learn from these
  Some funny ones

\subsection{Course blocks}

40\% of sessions have at least one block-click
Average of 4.94 block-clicks per session

\subsection{Social}

90 users have linked Skedge to Facebook
Since March 1st,
4,000+ visits (200 visits/day)
~60\% of visits to /social were returning visitors
90 overlays onto friends’ schedules
10 clicks to Facebook profiles :(
- get stats from the fb dashboard

\subsection{Conclusion}

Success! Considering skedge is OPTIONAL.
+ course blocks (obv usecase, can't click)
+ exports (not supported by thing)
+ mobile

\section{Navigations-per-add}

\subsection{Definitions}

A navigation is defined as
a search, or
a click on an instructor’s name, or
a click on a crosslisted or prerequisite course link

The navigations-per-{add, bookmark} measure is
the number of navigations a user took (within one session) until a course was {added, bookmarked}

\subsection{Trends}

%figs

\subsection{Breaking them apart}

  behavioral patterns
  Direct search for specific course
  Discovery, browsing, exploring

  \subsubsection{Direct searches}

  \subsubsection{Browse}

  % STAT: Find major of users, find how many non-major courses they found

\subsection{Conclusion}

%figs

Effective++

\section{Users' search types over time}

\subsection{Definitions}

Points for search by (omits number and dept.):

description
credits
crosslisted
CRN
instructor
title
year
term
‘random’
upper-level writing

“CSC” → 0
“MTH 165” → 0
“taught by hema” → 1   ✓    (2 searches) 
“random mur 1-2 credits” → 2   ✓    (1 search)

\subsection{Trends}

% fig

First increase (60.5% of users)
Median: 2 searches
Average: 4.23 searches
(Starting at 1 counts as an increase value of 0)

Second increase (7.9% of users)
Median: 8 searches
Average: 17.52 searches

\subsection{Conclusion}

DSL++