### 2016-04-09 conclusion.tex

\chapter{Conclusions}

\section{Proposal to the University}

\section{Resources}

\subsection*{Source code}

The source code for Skedge is available online under an open source license:\\
\url{https://github.com/RocHack/skedge}.

\subsection*{Live site}

\noindent The site can be found at: \url{http://skedgeur.com}.