### 2016-04-10 introduction.tex

\chapter{Introduction}

This paper will begin by

\section{Overview of CDCS}

CDCS is the University's official tool for

\label{fig:cdcs-index}

\subsection{``Better CDCS''}

%%%

\section{Overview of Skedge}

Skedge is a website I developed in 2014 and have been maintaining and developing since.

Bookmarks

Students, parents, department coordinators, and faculty can all benefit from such tool improvements.

\ref{fig:sk-index}

\begin{figure}[ht]
    \centering
        \begin{subfigure}[h]{14cm}
            \centering
            \fbox{
                \includegraphics[width=1.00\textwidth]{images/cdcs/index}
            }
            \caption{CDCS, with the search query {\tt csc}}
            \label{fig:cdcs-index}
        \end{subfigure}\\
        \vspace{10pt}\\
        \begin{subfigure}[h]{14cm}
            \centering
            \fbox{
                \includegraphics[width=1.00\textwidth]{images/cdcs/better}
            }
            \caption{Better CDCS, a separate browser extension that embeds buttons into the CDCS course results interface, allowing users to add courses to a locally-stored schedule}
            \label{fig:cdcs-better}
        \end{subfigure}
    \caption{CDCS and Better CDCS in their current states}
\end{figure}

\begin{figure}[ht]
    \centering
    \fbox{
        \includegraphics[width=1.00\textwidth]{images/skedge/index}
    }
    \caption{Skedge, with the search query {\tt csc} and the user's current schedule on the right}
    \label{fig:sk-index}
\end{figure}