### 2016-04-10 tech.tex

\chapter{Technical overview}

\section{Architecture}

Skedge's infrastructure is built

\emph{nginx}, \emph{unicorn}, \emph{Ruby on Rails}, \emph{PostgreSQL}, \emph{React.js}, \emph{Ahoy}, and Google Analytics.

\section{Data storage}

cookies, likes not on fb

\section{Usage data collection}