### 2016-04-11 conclusion.tex
Small fixes

\chapter{Conclusions}

\section{Looking forward}

\subsection{Features}

\subsection{Analytics}

%%%

\section{Proposal to the University}

%%%

\section{Acknowledgments}

%%%

\section{Resources}

\subsection*{Source code}

The source code for Skedge is available online under an open source license:\\
\url{https://github.com/RocHack/skedge}.

\subsection*{Live site}

\noindent The site can be found at \url{http://skedgeur.com} or \url{http://skedge.org} (very recently acquired, simply redirects to the former).