### 2016-04-11 tech.tex

\chapter{Technical overview}

\section{Architecture}

\emph{nginx}, \emph{unicorn}, \emph{Ruby on Rails}, \emph{PostgreSQL}, \emph{React.js},

\section{Data storage}

cookies, likes not on fb

\section{Usage data collection}

anonymous, \emph{Ahoy} and Google Analytics.