### 2016-04-12 tech.tex

\chapter{Technical overview}

\section{Architecture}

\emph{nginx}, \emph{unicorn}, \emph{Ruby on Rails}, \emph{PostgreSQL}, \emph{React.js},

\section{Data storage}

cookies, likes not on fb

\section{Usage data collection}

As the saying goes, ``if it's not monitored, it's not in production'' \cite{monitored} (how else would we know that it was?) anonymous, \emph{Ahoy} and Google Analytics.