### 2016-04-13 analytics.tex
Cheeseball

\chapter{Data analytics}

In this chapter I propose three hypotheses, each to be supported by analysis in the coming sections. The hypotheses are:

\begin{enumerate}
  \item Skedge's differences from and additions to CDCS are \textbf{usable and have real need}.

  \item Skedge’s \emph{navigations-per-add} demonstrate its \textbf{effectiveness} in the user search paradigms of \textbf{a) direct search} and \textbf{b) exploratory search}.

  \item \textbf{SkedgeQL is user-friendly}; users learn how to make more sophisticated search queries over time \textbf{simply by using it}.
\end{enumerate}

\input{analytics/usage}
\clearpage

\input{analytics/npa}
\clearpage

\input{analytics/dsl}
\clearpage