### 2016-04-13 tech.tex
Fix analytics intro

\chapter{Technical overview}

\section{System architecture}

\emph{nginx}, \emph{unicorn}, \emph{Ruby on Rails}, \emph{PostgreSQL}, \emph{React.js},

\section{Data storage}

users are lightweight and while persistent on the server, they rely on browser cookies for a client to access the back-end user

cookies, likes not on fb

\section{Usage data collection}

As the saying goes, ``if it's not monitored, it's not in production'' \cite{monitored} (how else would we know that it was?) anonymous, \emph{Ahoy} and Google Analytics.

A session is defined by a period of activity expiring after 30 minutes of inactivity