### 2016-04-13 paper.tex

\documentclass[titlepage]{report}

\usepackage{setspace}
\usepackage[comma,authoryear]{natbib}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}   
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{float}
\usepackage{hyperref}
\usepackage{listings}
\usepackage[labelfont=bf]{caption}
\usepackage[margin=3.5cm]{geometry}

\lstset{
	basicstyle=\footnotesize\ttfamily,
}

\title{
\vspace{60pt}\\
\huge \bfseries Skedge \\
\vspace{10pt}
\Large
Smarter course scheduling for our University
}

\author{
	Dan Hassin\\
    \vspace{5pt}\\
    Supervised by\\
    Professor Philip Guo\\
    \vspace{2pt}\\
    Department of Computer Science\\
    University of Rochester\\
    Rochester, New York\\
}

\date{April 13, 2016\\
    \vspace{80pt}\\
    submitted in partial fulfillment of\\
    the requirements for the degree\\
    \emph{honors bachelor of science},\\
    \vspace{40pt}\\
    and as an open letter addressed to\\
    the information technology staff\\
    at the university of rochester
    \vspace{-80pt}
}

\begin{document}

\maketitle

%%%%%%%%%%%

\onehalfspacing

\pagenumbering{gobble}
\setcounter{tocdepth}{1}
\tableofcontents

\listoffigures

\clearpage

%%%%%%%%%%%%%%%%%%

\doublespacing

\addcontentsline{toc}{chapter}{Abstract}

\begin{abstract}

\thispagestyle{plain}
\pagenumbering{roman}

In this paper I present Skedge, a web application for students to comfortably and effectively engage with the University's course catalog. Skedge matches and surpasses the capabilities of the existing University tool for this purpose, ``Course Description / Course Schedule'' (CDCS) and presents its information in a more visually appealing way. As a result, Skedge boasts strong user-retention rates, long session durations, and high student adoption despite having virtually no advertisement. Through collected usage data, I demonstrate that \textbf{a)} Skedge's differences from and additions to CDCS are usable and have real-world need, \textbf{b)} the three major use-cases associated with course scheduling---direct, exploratory, and peer-guided search---are effectively accommodated by Skedge, and \textbf{c)} Skedge's novel search mechanism is user-friendly and self-teaches to users over time.

\end{abstract}

%%%%%%%%%%%%%%%

\pagenumbering{arabic}

\setlength{\skip\footins}{0.75cm}

\include{introduction}
\clearpage

\include{design}
\clearpage

\include{tech}
\clearpage

\include{analytics}
\clearpage

\include{conclusion}
\clearpage

% References

\singlespacing

% Bibtex commands go here

\begin{spacing}{0.9}

\bibliographystyle{ieeetr}
\bibliography{bibliography}

\end{spacing}

\addcontentsline{toc}{chapter}{Bibliography}

\end{document}